---
title: "Random Forest Model Results"
author: "Storm Pino"
format: docx
---

## Random Forest Model Results (Storm Pino)

This section reports the Random Forest results for the Google Play Store app dataset. The goal was to predict an app’s outcome class using app characteristics.

The response variable was `Outcome_Class`, where:

- 1 = bad
- 2 = decent
- 3 = successful

The classes were defined using rating and install count:

- Class 3: Installs >= 1,000,000 and Rating >= 4.2
- Class 2: Installs >= 100,000 and Rating >= 3.5, but not enough to be Class 3
- Class 1: All remaining apps

The Random Forest model formula was:

Outcome_Class ~ Reviews + Size_MB + Price + Category + Type + Content_Rating

## Data Preparation

```{r, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(randomForest)
library(ggplot2)

df <- read_csv("data/processed/cleaned_data.csv", show_col_types = FALSE)

df <- df %>%
  rename(Price = `Price_$`)

df$Category <- as.factor(df$Category)
df$Type <- as.factor(df$Type)
df$Content_Rating <- as.factor(df$Content_Rating)

df <- df %>%
  mutate(
    Outcome_Class = case_when(
      Installs >= 1000000 & Rating >= 4.2 ~ 3,
      Installs >= 100000 & Rating >= 3.5 ~ 2,
      TRUE ~ 1
    )
  )

df$Outcome_Class <- as.factor(df$Outcome_Class)

table(df$Outcome_Class)
```

## Random Forest Method

Random Forest was used because it can handle nonlinear relationships and interactions between predictors. It also reduces the instability of a single decision tree by building many trees and combining their predictions.

A limitation of Random Forest is that it is less interpretable than simpler models. It can show which variables are important, but it does not give simple coefficients like linear or logistic regression.

## 10 Train/Test Splits

```{r, message=FALSE, warning=FALSE}
accuracy <- rep(0, 10)

for (i in 1:10) {
  
  set.seed(i)
  
  train_rows <- sample(1:nrow(df), size = 0.8 * nrow(df))
  
  train <- df[train_rows, ]
  test <- df[-train_rows, ]
  
  rf_model <- randomForest(
    Outcome_Class ~ Reviews + Size_MB + Price + Category + Type + Content_Rating,
    data = train,
    ntree = 1000,
    importance = TRUE
  )
  
  pred <- predict(rf_model, newdata = test)
  
  accuracy[i] <- mean(pred == test$Outcome_Class)
}

accuracy
mean_accuracy <- mean(accuracy)
mean_accuracy
```

## One Train/Test Split for Confusion Matrix

```{r, message=FALSE, warning=FALSE}
set.seed(123)

train_rows <- sample(1:nrow(df), size = 0.8 * nrow(df))

train <- df[train_rows, ]
test <- df[-train_rows, ]

rf_model <- randomForest(
  Outcome_Class ~ Reviews + Size_MB + Price + Category + Type + Content_Rating,
  data = train,
  ntree = 1000,
  importance = TRUE
)

pred <- predict(rf_model, newdata = test)

confusion_matrix <- table(Predicted = pred, Actual = test$Outcome_Class)
confusion_matrix

single_accuracy <- mean(pred == test$Outcome_Class)
single_accuracy
```

## Variable Importance

```{r, message=FALSE, warning=FALSE}
importance(rf_model)
varImpPlot(rf_model)
```

## Top Successful Categories

```{r, message=FALSE, warning=FALSE}
category_success <- df %>%
  group_by(Category) %>%
  summarise(
    total_apps = n(),
    successful_apps = sum(Outcome_Class == 3),
    success_rate = successful_apps / total_apps
  ) %>%
  filter(total_apps >= 20) %>%
  arrange(desc(success_rate))

head(category_success, 3)
```

```{r, message=FALSE, warning=FALSE}
ggplot(category_success, aes(x = reorder(Category, success_rate), y = success_rate)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Success Rate by App Category",
    x = "Category",
    y = "Success Rate"
  )
```

## Results

The Random Forest model achieved moderate predictive performance. Across 10 different train/test splits, the model’s average accuracy was calculated using the testing data from each split.

For one train/test split, the model produced an accuracy of about 62.5 percent. This means the model correctly classified about 62.5 percent of the apps in the testing set.

The model had difficulty predicting the higher outcome classes. This likely happened because the outcome classes were not evenly balanced. Since more apps were in Class 1, the model tended to predict the majority class more often.

## Important Predictors

The most important predictor was `Reviews`. Other important predictors were `Category`, `Price`, and `Content_Rating`. The variable `Size_MB` had little to no effect in the model.

This suggests that user engagement and app type were more useful for predicting app outcome than app size.

## Category Success Rates

We also examined which app categories had the highest success rates. Success rate was calculated as the number of Class 3 apps divided by the total number of apps in that category.

This helps answer whether certain categories are more likely to produce successful apps.

## Conclusion

The Random Forest model gave moderate predictive performance. It did not perfectly classify all app outcome levels, but it identified useful patterns in the data.

Overall, `Reviews`, `Category`, and `Price` were the strongest predictors of app outcome. App size was not useful for prediction.
